### 2018-03-16 paper/1-intro.tex
More changes

\item We show that we can instantiate
our abstractions to describe the essence of a variety of different
real-life build systems, including \Make, \Shake, \Bazel, and \Excel, each in
a dozen lines of code or so (\S\ref{sec-implementations}).
Doing this in a single setting allows
the differences and similarities between these huge systems to be
brought out clearly.

\item Moreover, we can readily mix the ingredients differently to produce the first
  build system that supports both \emph{dynamic dependencies}
  and \emph{cloud build} (\S\ref{sec-implementation-cloud-shake}).

@@

Papers about ``frameworks'' are often fuzzy.  This one is not: all our
abstractions are defined in Haskell, and we have (freely-available)
executable models of all the build systems we describe.  An unusual
feature is that we include \Excel in our line-up because, looked at
in the right way, it certainly is a build system.

>>>>

\item We show that we can instantiate
our abstractions to describe the essence of a variety of different
real-life build systems, including \Make, \Shake, \Bazel, and \Excel, each in
a dozen lines of code or so (\S\ref{sec-implementations}).
Doing this modelling in a single setting allows
the differences and similarities between these huge systems to be
brought out clearly.

\item Moreover, we can readily remix the ingredients to produce the first
  build system that supports both \emph{dynamic dependencies}
  and \emph{cloud build} (\S\ref{sec-implementation-cloud-shake}).

@@

Papers about ``frameworks'' are often fuzzy.  This one is not: all our
abstractions are defined in Haskell, and we have (freely-available)
executable models of all the build systems we describe.

### 2018-03-16 paper/2-background.tex
More changes

Build systems automate the execution of simple repeatable tasks for individual
users, as well as for large organisations. In this section we explore the design
space of build systems, using four concrete examples:
\Make~\cite{feldman1979make}, \Shake~\cite{mitchell2012shake},
\Bazel~\cite{bazel}, and \Excel~\cite{advanced_excel}\footnote{\Excel appears
very different to the others but, seen through the lens of this paper, it is
very close indeed.}.
We have carefully chosen these four to illustrate the various axes on which build systems
differ; we discuss many other notable examples
of build systems, and their relationships,
in~\S\ref{sec-related}-\S\ref{sec-conclusions}.

>>>>

Build systems automate the execution of simple repeatable tasks for individual
users, as well as for large organisations. In this section we explore the design
space of build systems, using four concrete examples:
\Make~\cite{feldman1979make}, \Shake~\cite{mitchell2012shake},
\Bazel~\cite{bazel}, and \Excel~\cite{advanced_excel}\footnote{\Excel appears
very different to the others but, seen through the lens of this paper, it is
very close indeed.}.
We have carefully chosen these four to illustrate the various axes on which build systems
differ; we discuss many other notable examples
of build systems, and their relationships,
in~\S\ref{sec-related}.

### 2018-03-16 paper/4-build.tex
More changes

As per the definition of minimality~\ref{def-minimal}, a minimal build
system must \underline{rebuild only out-of-date keys} and at most once. The only
way to achieve the ``at most once'' requirement while producing a correct build
result (\S\ref{sec-build-correctness}) is to \underline{build all keys in the
order that respects their dependencies}.

@@

This approach requires a way to suspend a running task, which can be done with cheap
green threads and blocking (the original approach of \Shake) or with a
continuation-passing style (what \Shake does now). An alternative approach to
suspending a task is to abort it and restart it again later, at the cost
of doing additional work.

@@

When using a dirty bit, it is necessary to check all the dependencies of a key. For applicative build systems that list is simple to obtain, but for monadic build systems there is no general way to get all dependencies. Instead \Excel computes a \textit{static approximation} of the dependencies. For applicative rules that approximation is correct. For functions such as \textsf{IF} it marks the cell dirty if \textit{any} potential dependency has changed, even on the untaken \textif{if} branch. For functions such as \textsf{INDIRECT} whose dependencies cannot be guessed, it conservatively assumes the dependencies have always changed.

With a dirty bit it is simple to achieve minimality. However, to achieve early cutoff
(\S\ref{sec-background-shake}) it would be important to not set the dirty bit
after a computation that did not change the value. \Excel could use this
approach, but does not. In contrast, \Make cannot implement early cutoff -- to
do so it would have to mark the node clean (so it would not rebuild in the next
run) and at the same time not mark the things it depends on dirty -- an
impossible task with only the ability to update to the latest modification time.

>>>>

As per the definition of minimality~\ref{def-minimal}, a minimal build
system must \underline{rebuild only out-of-date keys} and at most once. The only
way to achieve the ``at most once'' requirement while producing a correct build
result (\S\ref{sec-build-correctness}) is to \underline{build all keys in an
order that respects their dependencies}.

@@

This approach requires that a task may be started, then during that execution
another task will have to be run. Assuming a an IO-driven task structure,
that requires suspend a running task, which can be done with cheap
green threads and blocking (the original approach of \Shake) or with a
continuation-passing style (what \Shake does now). An alternative approach to
suspending a task is to abort it and restart it again later, at the cost
of doing additional work.

@@

When using a dirty bit, it is necessary to check all the dependencies of a key. For applicative build systems that list is simple to obtain, but for monadic build systems there is no general way to get all dependencies. Instead \Excel computes a \textit{static approximation} of the dependencies. For applicative rules that approximation is correct. For functions such as \textsf{IF} it marks the cell dirty if \textit{any} potential dependency has changed, even on the untaken \textsf{IF} branch. For functions such as \textsf{INDIRECT} whose dependencies cannot be guessed, it conservatively assumes the dependencies have always changed.

With a dirty bit it is simple to achieve minimality. However, to achieve early cutoff
(\S\ref{sec-background-shake}) it would be important to not set the dirty bit
after a computation that did not change the value. \Excel could use this
approach, but does not. In contrast, \Make cannot implement early cutoff nicely -- to
do so it would have to mark the node clean (so it would not rebuild in the next
run) and at the same time not mark the things it depends on dirty -- an
impossible task with only the ability to update to the latest modification time.
It can approximate early cutoff by not touching the result file, and not marking it clean,
but then it reruns in every subsequent build.
